HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-05-22 tmlr_submission/hermes_paper_revision/hermes_paper.tex
add new result for tmlr revision

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citet{hyndman2020}. \textcolor{orange}{As deep learning apporaches, two recurrent neural networks are proposed. They are both composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52. The first one (called \textit{lstm}) has only access to the main signal while the second one (called \textit{lstm-ws}) has access to the main and the external signal.} \textcolor{gray}{As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.}

@@

\textcolor{orange}{Considering the new HERMES approaches, all the different versions show consequent improvements regarding their per-time-series predictors in terms of MASE. \textit{Hermes-tbats} largely outperforms the two other declinations and all the benchmarks on the fashion dataset. This result highlight to important point of the proposed hybrid model: i) The final accuracy of the model is strongly impacted by the choice of per-time-series predictor ii) If the per-time-series predictor is well chosen regarding the nature of the time series, the resulting HERMES model can show impressive results and outperformance state-of-the-art approaches.}
\textcolor{gray}{Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \citet{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.}

@@

\begin{table}
  \caption{Results summary on the 10000ts Fashion dataset. For each metric, the average on all our time series is computed. For approaches using neural networks, 10 models are trained with different seeds. The mean and the standard deviation of the 10 results are displayed.}
  \centering
  \begin{tabular}{l||lllll|lllll}
   &&\multicolumn{3}{c}{\textbf{MASE $\downarrow$}} &&& \multicolumn{3}{c}{\textbf{ACCURACY $\uparrow$}}&\\
    &&  \textit{mean}  && \textit{std} &&&  \textit{mean}  && \textit{std}& \\
	 \hline
	 &&&&&&&&&&\\
     \textit{snaive} && 0.881 && - &&& 0.357 && - &\\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.0.845} && \textcolor{orange}{-} &&& \textcolor{orange}{0.463} && \textcolor{orange}{-}\\
     \textit{arima} && 0.826 && -&&& 0.464 && - & \\
     \textit{ets} && 0.807 && -&&& 0.449 && - & \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.786} && \textcolor{orange}{-} &&& \textcolor{orange}{0.485} && \textcolor{orange}{-}\\
     \textit{stlm} && 0.770 && -&&& 0.482 && - & \\
     \textit{hermes-ets-ws} && 0.769 && 0.005 &&& 0.501 && 0.007 &\\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.764} && \textcolor{orange}{0.003} &&& \textcolor{orange}{0.497} && \textcolor{orange}{0.005}\\
     \textcolor{orange}{\textit{hermes-thetam-ws}} && \textcolor{orange}{0.760} && \textcolor{orange}{0.004} &&& \textcolor{orange}{\textbf{0.520}} && \textcolor{orange}{0.01}\\
     \textit{hermes-ets} && 0.758 && 0.001 &&& 0.490 && 0.006 &\\
     \textit{tbats} && 0.745 && -&&& 0.453 && - & \\
     \textit{lstm-ws} && 0.728 && 0.004 &&& 0.500 && 0.008 &\\
     \textit{lstm} && 0.724 && 0.003 &&& 0.498 && 0.007 &\\
     \textit{hermes-tbats} && 0.715 && 0.002 &&& 0.488 && 0.008 &\\
     \textbf{\textit{hermes-tbats-ws}} && \textbf{0.712} && 0.004 &&& 0.510 && 0.005 &\\
  \end{tabular}
\label{tab:metricresults}
\end{table}

@@

\textbf{Size of the dataset. } In addition to the results on the whole fashion dataset, the robustness of the HERMES model is analyzed when it is trained on smaller datasets. Two experiments are performed on a sub sample of respectively 1000 and 100 randomly selected time series. Results are given in Table~\ref{tab:1000metricresults}. The hybrid framework \textit{hermes-tbats} achieves the best performance in terms of global accuracy on both datasets. We can note that the accuracy of the full neural network \textit{lstm} decreases when the dataset size decreases. On the small dataset of 100 time series, a local statistical model like \textit{tbats} or \textit{stlm} largely outperforms the \textit{lstm}. Providing sharp predictions from scratch is a complex task and high-dimensional recurrent neural networks require large amounts of data to do so. By contrast, the HERMES approach can rely on its first statistical part and consequently needs less data to be trained and to obtain interesting performance. \textcolor{orange}{We can nevertheless note that the gain brought by the error-corrector recurrent model decreases as the size of the dataset diminish.}

\begin{table}
  \caption{Results summary on the 1000 time series and 100 time series Fashion dataset. The MASE average on all the time series is computed. For the two approaches using a neural network, 10 models with different seeds are trained. the mean and the standard deviation of the 10 results are displayed.}
 \centering
 \resizebox{0.8\textwidth}{!}{ 
  \begin{tabular}{l||llll}
    \multicolumn{4}{c}{1000 time series Fashion dataset}\vspace{0.5cm} \\
    &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	 \hline
	 &&& \\
     \textit{snaive} && 0.871 & - \\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.837} & \textcolor{orange}{-} \\
     \textit{arima} && 0.821 & - \\
     \textcolor{orange}{\textit{ets}} && \textcolor{orange}{0.801} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.788} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{hermes-ets}} && \textcolor{orange}{0.768} & \textcolor{orange}{0.005} \\
     \textit{stlm} && 0.765 & - \\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.764} & \textcolor{orange}{0.004} \\
     \textit{lstm} && 0.740 & 0.007 \\
     \textit{tbats} && 0.734 & - \\
     \textbf{\textit{hermes-tbats}} && \textbf{0.719} & 0.002 \\
  \end{tabular}\hspace{1cm}
\vspace{.2cm}

  \begin{tabular}{l||llll}
   \multicolumn{4}{c}{100 ts Fashion dataset}\vspace{0.5cm} \\
   &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	\hline
	 &&& \\
     \textit{snaive} && 0.876 & - \\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.822} & \textcolor{orange}{-} \\
     \textit{arima} && 0.814 & - \\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.808} & \textcolor{orange}{0.006} \\
     \textcolor{orange}{\textit{hermes-ets}} && \textcolor{orange}{0.792} & \textcolor{orange}{0.003} \\
     \textcolor{orange}{\textit{ets}} && \textcolor{orange}{0.786} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.767} & \textcolor{orange}{-} \\
     \textit{lstm} && 0.767 & 0.045 \\
     \textit{stlm} && 0.742 & - \\
     \textit{tbats} && 0.745 & - \\
     \textbf{\textit{hermes-tbats}} && \textbf{0.739} & 0.003 \\
  \end{tabular}
 }
\label{tab:1000metricresults}
\end{table}

>>>>

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citet{hyndman2020}. \textcolor{orange}{To complete the statistical benchmarks, the Prophet model introduced in \citet{Taylor2017} is proposed and implemented with the available package bearing the same name \texttt{prophet}. As deep learning apporaches, Three models based on recurrent neural networks are proposed. They are called \textit{lstm}, \textit{lstm-ws} and \textit{deepar}. Concerning \textit{lstm} and \textit{lstm-ws}, they are both composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52. The first one (\textit{lstm}) has only access to the main signal while the second one (\textit{lstm-ws}) has access to the main and the external signal. The last approach is the recent state-of-the-art model called DeepAR and introduced in \citet{salinas2020}. The python package \texttt{GluonTS} \citep{Alexandrov2020} was used to train the DeepAR approach on the fashion dataset and a grisearch on the main parameters have been done.} \textcolor{gray}{As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.}

@@

\textcolor{orange}{Considering the new HERMES approaches, all the different versions show consequent improvements regarding their per-time-series predictors in terms of MASE. \textit{Hermes-tbats} largely outperforms the two other declinations and all the benchmarks on the fashion dataset. This result highlight to important point of the proposed hybrid model: i) The final accuracy of the model is strongly impacted by the choice of per-time-series predictor ii) If the per-time-series predictor is well chosen regarding the nature of the time series, the resulting HERMES model can show impressive results and outperformance state-of-the-art approaches such as DeepAR.}
\textcolor{gray}{Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \citet{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.}

@@

\begin{table}
  \caption{Results summary on the 10000ts Fashion dataset. For each metric, the average on all our time series is computed. For approaches using neural networks, 10 models are trained with different seeds. The mean and the standard deviation of the 10 results are displayed.}
  \centering
  \begin{tabular}{l||lllll|lllll}
   &&\multicolumn{3}{c}{\textbf{MASE $\downarrow$}} &&& \multicolumn{3}{c}{\textbf{ACCURACY $\uparrow$}}&\\
    &&  \textit{mean}  && \textit{std} &&&  \textit{mean}  && \textit{std}& \\
	 \hline
	 &&&&&&&&&&\\
     \textit{snaive} && 0.881 && - &&& 0.357 && - &\\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.845} && \textcolor{orange}{-} &&& \textcolor{orange}{0.463} && \textcolor{orange}{-}\\
     \textit{arima} && 0.826 && -&&& 0.464 && - & \\
     \textit{ets} && 0.807 && -&&& 0.449 && - & \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.786} && \textcolor{orange}{-} &&& \textcolor{orange}{0.485} && \textcolor{orange}{-}\\
     \textit{stlm} && 0.770 && -&&& 0.482 && - & \\
     \textit{hermes-ets-ws} && 0.769 && 0.005 &&& 0.501 && 0.007 &\\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.764} && \textcolor{orange}{0.003} &&& \textcolor{orange}{0.497} && \textcolor{orange}{0.005}\\
     \textcolor{orange}{\textit{hermes-thetam-ws}} && \textcolor{orange}{0.760} && \textcolor{orange}{0.004} &&& \textcolor{orange}{\textbf{0.520}} && \textcolor{orange}{0.010}\\
     \textit{hermes-ets} && 0.758 && 0.001 &&& 0.490 && 0.006 &\\
     \textcolor{orange}{\textit{deepar}} && \textcolor{orange}{0.752} && \textcolor{orange}{0.018} &&& \textcolor{orange}{0.459} && \textcolor{orange}{0.015}\\
     \textit{tbats} && 0.745 && -&&& 0.453 && - & \\
     \textit{lstm-ws} && 0.728 && 0.004 &&& 0.500 && 0.008 &\\
     \textit{lstm} && 0.724 && 0.003 &&& 0.498 && 0.007 &\\
     \textit{hermes-tbats} && 0.715 && 0.002 &&& 0.488 && 0.008 &\\
     \textbf{\textit{hermes-tbats-ws}} && \textbf{0.712} && 0.004 &&& 0.510 && 0.005 &\\
  \end{tabular}
\label{tab:metricresults}
\end{table}

@@

\textbf{Size of the dataset. } In addition to the results on the whole fashion dataset, the robustness of the HERMES model is analyzed when it is trained on smaller datasets. Two experiments are performed on a sub sample of respectively 1000 and 100 randomly selected time series. Results are given in Table~\ref{tab:1000metricresults}. The hybrid framework \textit{hermes-tbats} achieves the best performance in terms of global accuracy on both datasets. We can note that the accuracy of the full neural network \textit{lstm} decreases when the dataset size decreases. On the small dataset of 100 time series, a local statistical model like \textit{tbats} or \textit{stlm} largely outperforms \textcolor{orange}{deep-model-based approach such as \textit{lstm} of \textit{deepar}}. In fact, providing sharp predictions from scratch is a complex task and high-dimensional recurrent neural networks require large amounts of data to do so. By contrast, the HERMES approach can rely on its first statistical part and consequently needs less data to be trained and to obtain interesting performance. \textcolor{orange}{We can nevertheless note that the gain brought by the error-corrector recurrent model decreases as the size of the dataset diminish.}

\begin{table}
  \caption{Results summary on the 1000 time series and 100 time series Fashion dataset. The MASE average on all the time series is computed. For the two approaches using a neural network, 10 models with different seeds are trained. the mean and the standard deviation of the 10 results are displayed.}
 \centering
 \resizebox{0.8\textwidth}{!}{ 
  \begin{tabular}{l||llll}
    \multicolumn{4}{c}{1000 time series Fashion dataset}\vspace{0.5cm} \\
    &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	 \hline
	 &&& \\
     \textit{snaive} && 0.871 & - \\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.837} & \textcolor{orange}{-} \\
     \textit{arima} && 0.821 & - \\
     \textcolor{orange}{\textit{ets}} && \textcolor{orange}{0.801} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.788} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{hermes-ets}} && \textcolor{orange}{0.768} & \textcolor{orange}{0.005} \\
     \textit{stlm} && 0.765 & - \\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.764} & \textcolor{orange}{0.004} \\
     \textit{lstm} && 0.740 & 0.007 \\
     \textcolor{orange}{\textit{deepar}} && \textcolor{orange}{0.738} & \textcolor{orange}{0.017} \\
     \textit{tbats} && 0.734 & - \\
     \textbf{\textit{hermes-tbats}} && \textbf{0.719} & 0.002 \\
  \end{tabular}\hspace{1cm}
\vspace{.2cm}

  \begin{tabular}{l||llll}
   \multicolumn{4}{c}{100 ts Fashion dataset}\vspace{0.5cm} \\
   &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	\hline
	 &&& \\
     \textit{snaive} && 0.876 & - \\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.822} & \textcolor{orange}{-} \\
     \textit{arima} && 0.814 & - \\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.808} & \textcolor{orange}{0.006} \\
     \textcolor{orange}{\textit{deepar}} && \textcolor{orange}{0.804} & \textcolor{orange}{0.024} \\
     \textcolor{orange}{\textit{hermes-ets}} && \textcolor{orange}{0.792} & \textcolor{orange}{0.003} \\
     \textcolor{orange}{\textit{ets}} && \textcolor{orange}{0.786} & \textcolor{orange}{-} \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.767} & \textcolor{orange}{-} \\
     \textit{lstm} && 0.767 & 0.045 \\
     \textit{stlm} && 0.742 & - \\
     \textit{tbats} && 0.745 & - \\
     \textbf{\textit{hermes-tbats}} && \textbf{0.739} & 0.003 \\
  \end{tabular}
 }
\label{tab:1000metricresults}
\end{table}
